---
title: 'Combined-Model Testing'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: html_document
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---

## Introduction

This file is used to test out the combined-model approach proposed in ADD DISSERTATION PROQUEST CITATION WHEN PUBLISHED.

## Preamble

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

First, let's load the necessary packages. I've also set this code chunk to not print warnings, as they are largely not helpful for the current example.

```{r libraries , warning = FALSE}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
library(rlang)      # For parse_expr() to parse data name in k-fold subsetting functions
library(msaenet)    # For the multi-step adaptive elastic net
```


Let's also load our test data, which will just be our matricized singular applied dataset.

```{r load synthetic applied dataset}
demo.data <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/TestingApplied_Xmtx.RData")
```

Let's also quickly combine the X and Y list elements into a single object for convenience with the subsetting function. We will put Y into the first column since it corresponds with a left-to-right reading of a regression model.

```{r combine X and Y into single object}
demo.data2 <- setDT(data.frame(cbind(demo.data[["Y"]] , demo.data[["X"]]
                                     )
                               )
                    )
```

## Loading custom functions

### Custom k-fold subsetting function

The chunk below takes a dataset and creates an ordinal variable with *k* levels, corresponding with each of the *k* test sets. The function takes as its arguments:

 * **data** : a dataset to be split into *k* groups
 * ***k*** : the number of folds for cross-validation
   * Defaults to 5 folds
   * Corresponds with the percentage of data to be used as a holdout/testing set in the following manner:
   $$
   \begin{equation}
        n_{test} = \text{
                $n_{full}$ $/$ $k$
        }
   \end{equation}
   $$
 * **seed** : a seed for fixing random processes
   * For this function, this is primarily associated with random and unordered assignment of the values of *k* across the full dataset
 * **list** : this argument tells the function if you want the resulting *k*-fold indexed data to be put into list format, with each list element corresponding with one fold *k*
   * This function defaults to *FALSE*, meaning the resulting data will be in matrix/tabular format and will contain a separate column for the index *k*
   * If set to *TRUE*, the resulting list will *not* contain the index variable for *k*, as the list elements themselves correspond with *k*
   * If set to *"traintest"* (in quotes!), the resulting list will instead contain two elements:
     * A training set containing data from *k* - 1 folds/subsets
     * A holdout/testing set containing data from the *k* fold/subset

```{r kfold subsetting function}
# k-fold subsetting function
kfold_subsetter <- function(data , y_col = 1 , 
                            x_cols = c(2:ncol(data)) , 
                            subset_col = (ncol(data) + 1) , 
                            k = 5 , seed = 7 , list = FALSE) {
        # check for string object in data argument
        if(is.character(data)) {
                data <- eval(parse_expr(data))
        }
        
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # check for 0 < k <= n/2
        if((k <= 0) | (k > (nrow(data) / 2) ) ) {
                stop("ERROR: number of folds 'k' must be greater than 0 and less than sample size divided by 2")
        }
        
        # determine number of subsets which contain an extra element
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
                
        newdata <- setDT(data.frame(newdata))
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                    X = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset < k) %>%
                        .[ , c(1:8) , with = F] %>%
                        as.matrix() , 
                    Y = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset < k) %>%
                        .[ , 1 , with = F] %>%
                        as.matrix() , 
                    X_Test = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset %in% k) %>%
                        .[ , c(1:8) , with = F] %>%
                        as.matrix() , 
                    Y_Test = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset %in% k) %>%
                        .[ , 1 , with = F] %>%
                        as.matrix() , 
                    Seed = seed , 
                    Subsets = newdata[ , "subset"]
                        #Train = subset(newdata , subset < k) , 
                        #Test = subset(newdata , subset %in% k) , 
                        #Seed = seed
                        )
                        )
        } else {
                newdata <- return(newdata)
        }
}



```

### Multi-Subsetting Wrapper

The following function serves as a wrapper around the original subsetting function which incorporates additional arguments. These arguments extend the original function to conduct multiple training-testing splits and store the resulting datasets in a singular overarching list.

The wrapper's arguments are as follows. The ellipses are arguments for the original subsetting function which are handled internally within the wrapper.

 * **data** : a dataset to be split into training-testing sets
 * **seed_multi** : a seed for generating the vector of seeds for each individual training-testing *k*-fold split
 * **num_splits** : the number of training-testing splits to be created
    * Defaults to 100
 * **test_percent** : the percentage (as a decimal) of the full data to be reserved for the holdout/testing set
    * Corresponds with the number of folds *k* for *k*-fold subsetting in the following manner:
   $$
   \begin{equation}
        test\_percent = \text{
                $1$ $/$ $k$
        }
   \end{equation}
   $$
    * Defaults to 20$\%$, which corresponds with 5-fold subsetting

Note that this wrapper defaults to setting "list" argument in *kfold_subsetter()* to "traintest." This decision was implemented to reduce issues with formatting an applied dataset for the final combined-model function. The subsetting function is relatively more flexible in terms of data inputs than the full combined-model function, therefore I wanted to streamline data formatting from that point forward when using the combined model in full.

```{r multi-kfold wrapper}
kfold_multi <- function(data , ... , seed_multi = 713 , 
                        num_splits = 100 , test_percent = .2) {
        # set the seed for generating individual kfold seeds
        set.seed(seed_multi)
        
        # turn data object name into character string for subsequent use
        data.name <- deparse(substitute(data))
        
        # generate random numbers equal to num_splits so that each
        # # individual run of the subsetter has a unique seed
        seeds <- sample(c(1:100000) , size = num_splits , replace = FALSE)
        
        # set value of k corresponding with test_percent
        if((test_percent >= 1) | (test_percent <= 0)) {
                stop("ERROR: 'test_percent' must be greater than and less than 1")
        }
        
        # set k for individual splits
        k <- 1 / test_percent
        
        # initialize and data.table of kfold arguments
        split_repped.dt <- setDT(as.data.frame(matrix(ncol = 1 , 
                                              nrow = num_splits)))
        
        # fill data.table with arguments for kfold function
        split_repped.dt[ , ':=' (data = data.name , 
                         k = k , 
                         seed = seeds , 
                         list = "traintest"
                         )
                 ]
        
        # remove blank column from initializing
        split_repped.dt <- split_repped.dt[ , !1 , with = F]
        
        # run subsetter for the desired number of splits
        full.data <- split_repped.dt %>%
                pmap(kfold_subsetter)
        
        # store single object of all training/testing splits
        return(full.data)
}
```

### Model-Application Function

Finally, let's load the function used to generate our the models from our lasso and elastic net adaptations.

```{r hqreg model application function , echo = T}

# adaptive LAD lasso application function
hqmsa.sim.fnct <- function(data.list , 
                           method = c("msaenet" , 
                                      "quantile" , "LAD" , "huber") , 
                           tau = 0.5 , gamma = 1.345 , alpha = 0.5 , nsteps = 10L) {
       # Store training X and Y to temporary objects
       X_train <- data.list[["X"]]
       Y_train <- data.list[["Y"]]
       
       if(method == "LAD") {
               method <- "quantile"
       }
       
       # If applicable, store holdout/testing X and Y
       if(!is.null(data.list[["X_test"]]) & 
          !is.null(data.list[["Y_test"]])) {
               X_test <- data.list[["X_test"]]
               Y_test <- data.list[["Y_test"]]
       } else {
               X_test <- NULL
               Y_test <- NULL
       }
        
       # lambdas to try for regularization
       lambda.try <- seq(log(1400) , log(0.01) , length.out = 100)
       lambda.try <- exp(lambda.try)
       
       # set a timer start point
       start <- Sys.time()
       
       # cross-validated selection of adaptive lasso
       # # tuning hyperparameter nu/gamma
       
       # # select ridge coefs for weighting
       ridge.model <- cv.glmnet(x = X_train , y = Y_train , 
                                lambda = lambda.try , alpha = 0)
       lambda.ridge.opt <- ridge.model$lambda.min
       best.ridge.coefs <- predict(ridge.model , 
                                           type = "coefficients" ,
                                           s = lambda.ridge.opt)[-1]

       
       # # grid of nu/gamma values to try for cross-validation
       nu.try <- exp(seq(log(0.01) , log(10) , length.out = 100))
       
       # # initialize full list of LAD lasso results from each nu/gamma
       hqmsa.nu.cv.full <- list()
       
       # # initialize matrices of metrics and minimizing results
       hqmsa.nu.cv.lambda <- numeric()
       hqmsa.nu.cv.mse <- numeric()
       hqmsa.nu.cv.msesd <- numeric()
       hqmsa.nu.cv.coefs <- list()
       
       # # Loop over nu/gamma values for CV, 
       # # # storing minimizing lambda within each nu/gamma
       if(method == "msaenet") {
               for(i in 1:length(nu.try)) {
                       #single adaptive lasso run with ridge weighting and nu = 1
                       hqmsa.nu.cv.full[[i]] <- msaenet(x = X_train , 
                                                    y = Y_train , 
                                                    family = "gaussian" , 
                                                    init = "ridge" ,
                                                    alphas = 0.5 , 
                                                    tune = "cv" , 
                                                    nfolds = 5L , 
                                                    rule = "lambda.min" , 
                                                    nsteps = nsteps , 
                                                    tune.nsteps = "max" , 
                                                    scale = nu.try[i])
                       
                       hqmsa.nu.cv.lambda[i] <-
                               hqmsa.nu.cv.full[[i]]$best.lambdas[[nsteps + 1]]
                       
                       hqmsa.nu.cv.coefs[[i]] <- c(NA , coef(hqmsa.nu.cv.full[[i]]))
                               
                       
                       hqmsa.nu.cv.mse[i] <- min(hqmsa.nu.cv.full[[i]]$step.criterion[[nsteps + 1]])
                       }
       } else {
               for(i in 1:length(nu.try)) {
                       invisible(capture.output(
                               hqmsa.nu.cv.full[[i]] <- 
                                       cv.hqreg(X = X_train , 
                                                y = Y_train , 
                                                method = method , 
                                                tau = tau , 
                                                gamma = gamma , 
                                                lambda = lambda.try ,
                                                alpha = alpha , 
                                                preprocess =
                                                        "standardize" , 
                                                screen = "ASR" , 
                                                penalty.factor = 
                                                        1 / abs(best.ridge.coefs) ^ nu.try[i] , 
                                                FUN = "hqreg" , 
                                                type.measure = "mse"
                                                )
                                        )
                               )
                       hqmsa.nu.cv.mse[i] <-
                               min(hqmsa.nu.cv.full[[i]]$cve)
                       hqmsa.nu.cv.msesd[i] <-
                               hqmsa.nu.cv.full[[i]]$cvse[
                         which.min(hqmsa.nu.cv.full[[i]]$cve)
                                                         ]
                       hqmsa.nu.cv.lambda[i] <-
                               hqmsa.nu.cv.full[[i]]$lambda.min
                       hqmsa.nu.cv.coefs[[i]] <- 
                               hqmsa.nu.cv.full[[i]]$fit$beta[ , 
                         which.min(hqmsa.nu.cv.full[[i]]$cve)
                                                             ]
                       }
       }

       
       #specify minimizing nu value and resulting model info
       nu.opt <- nu.try[which.min(hqmsa.nu.cv.mse)]
       lambda.opt <- 
               hqmsa.nu.cv.lambda[
                       which.min(hqmsa.nu.cv.mse)
                       ]
       weights.opt <- 1 / abs(best.ridge.coefs) ^ nu.opt
       hqmsa.coefs <-
               hqmsa.nu.cv.coefs[[
                       which.min(hqmsa.nu.cv.mse)
                       ]]
       hqmsa.mse.min <- min(hqmsa.nu.cv.mse)
       if(!is.null(hqmsa.nu.cv.msesd[1])) {
               hqmsa.mse.min.se <- hqmsa.nu.cv.msesd[
                       which.min(hqmsa.nu.cv.mse)
                       ]               
       }

       hqmsa.model.min <- 
               hqmsa.nu.cv.full[
                       which.min(hqmsa.nu.cv.mse)
                       ]
       n.coefs <- sum(hqmsa.coefs[-1] != 0)
       
       # calculate metrics using holdout data, if applicable
       if(!is.null(X_test) & !is.null(Y_test)) {
               # store n
               n <- nrow(data.list[["X_test"]])
               
               # calculate predicted values
               y.pred <- data.list[["X_test"]] %*% hqmsa.coefs[-1]
               if(!is.na(hqmsa.coefs[1])) {
                       y.pred <- y.pred + hqmsa.coefs[1]
               }
   
               # calculate residual
               resid <- y.pred - Y_test
               
               # square the residuals
               resid.sq <- resid ^ 2
               
               # sum the square of residuals
               sum.resid.sq <- sum(resid.sq)
               
               #calculate root mse
               mse <- sum.resid.sq / n
               
               # set endpoint for timer
               end <- Sys.time()
               
               # temporarily store time of current model
               time <- abs(as.numeric(difftime(start , 
                                               end , 
                                               units = "secs"
                                               )
                                      )
                           )
               
               # print the total runtime of the current model
               cat("time = " , time , "\n")
               
               # put conditions, model info, and metrics into list
               return(list(full.model = hqmsa.model.min ,
                           model.info = list(lambda = lambda.opt , 
                                             coefs = hqmsa.coefs , 
                                             weights = weights.opt
                                             ) , 
                           metrics = list(n.coefs = n.coefs , 
                                          runtime = time , 
                                          mse = mse
                                          )
                           )
                      )
       } else {
               # set endpoint for timer
               end <- Sys.time()
               
               # temporarily store time of current model
               time <- abs(as.numeric(difftime(start , 
                                               end , 
                                               units = "secs"
                                               )
                                      )
                           )
               
               # print the total runtime of the current model
               cat("time = " , time , " ;;; ")
               
               # put conditions, model info, and metrics into list
               return(list(full.model = hqmsa.model.min ,
                           model.info = list(lambda = lambda.opt , 
                                             coefs = hqmsa.coefs , 
                                             weights = weights.opt
                                             ) , 
                           metrics = list(n.coefs = n.coefs , 
                                          runtime = time
                                          )
                           )
                      )
       }
       
       

       

}
```

### Splitting the Data

Before we apply our models, let's create our data splits. We will use the multi-subsetting wrapper above to generate 100 random data splits. We will reserve 20$\%$ of the data for holdout/generating prediction error (aka, 5-fold subsets).

```{r splitting data}
demo.split <- list()

demo.split<- kfold_multi(demo.data2 , seed_multi = 999 , 
                         num_splits = 100 , test_percent = 0.2)
```

Here is a good place to stop and save the split data if you want store it for later use using the _saveRDS()_ function.

## Running our models

Now we can generate the models from our adaptations! I opted to generate the 100-model objects for each adaptation separately and then combined them. A little work with tools use throughout the different component demos could make a wrapper that generates all with one command and store them in a single objects. I will leave that as an exercise for anyone looking to practice function-writing. 

We are going to generate models from each of the following adaptations:

 * Adaptive LAD asso
 * Adaptive LAD elastic net, $\alpha = 0.5$
 * Adaptive Huber lasso, $M = 1.345$
 * Adaptive Huber elastic net, $M = 1.345, \alpha = 0.5$
 * Multi-step adaptive elastic net with $k = 3$ adaptive steps, $\alpha = 0.5$
 * Multi-step adaptive elastic net with $k = 5$ adaptive steps, $\alpha = 0.5$
 * Multi-step adaptive elastic net with $k = 10$ adaptive steps, $\alpha = 0.5$



### Adaptive LAD Lasso

```{r ladlasso}
#run across data subset
ladlasso.full <- demo.split %>%
        map(safely(hqmsa.sim.fnct) , method = "LAD" , alpha = 1)
```

```{r ladelnet5}
#run across data subset
ladelnet5.full <- demo.split %>%
        map(safely(hqmsa.sim.fnct) , method = "LAD" , alpha = 0.5)
```

```{r huberlasso}
#run across data subset
huberlasso.full <- demo.split[1:2] %>%
        map(safely(hqmsa.sim.fnct) , method = "huber" , alpha = 1)
```

```{r ladelnet5}
#run across data subset
huberelnet5.full <- demo.split %>%
        map(safely(hqmsa.sim.fnct) , method = "huber" , alpha = 0.5)
```

```{r msaenet k3}
#run across data subset
msaelnet.full <- demo.split %>%
        map(safely(hqmsa.sim.fnct) , method = "msaenet" , nsteps = 3L)
```
