---
title: "Scratch_Paper"
author: "Matt Multach"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    #toc: false
    #toc_depth: 4
    #toc_float: true
    #number_sections: false
bibliography: ../Component_Demos/Component_Refs.bib
---

# Preamble and Setup {.unlisted .unnumbered .tabset .tabset-fade .tabset-pills}



```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

<style>
    .ref {
        position: relative;
        vertical-align: baseline;
    }

    .refnum {
        position: relative;
        left: -1px;
        bottom: 1ex;
        font-family: Verdana, sans-serif;
        color: #005994;
        font-size: .7em;
        font-weight: bold;
        text-decoration: underline;
        cursor: pointer;        
    }

    .refbody {
        font-family: Verdana, sans-serif;
        font-size: .7em;
        line-height: 1.1;
        display: none;
        min-width: 20em;
        position: absolute;
        left: 25px;
        bottom: 5px ;
        border: 1px solid;
        padding: 5px;
        background-color: #fff;
        word-wrap: break-word;
        z-index: 9999;
        overflow: auto;
    }

</style>

<script>
    function footypop(id) {
        var el = document.getElementById(id) ;
  
        if (el.style.display == "none") {
            el.style.display = "block" ;
        } else {
            el.style.display = "none" ;
        }
        
    }
</script>



## Packages {.unnumbered}

<details>

```{r libraries , warning = FALSE , message = FALSE}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
library(rlang)      # For parse_expr() to parse data name in k-fold subsetting functions
library(msaenet)      # For the multi-step adaptive elastic net
library(gridExtra)    # For displaying grids of objects
library(grid)         # For creating a title for multiplot grid
library(ggplot2)      # For generating and manipulation graphical objects
```

</details>


  

## Testing collapsible notes {.unnumbered}

Does the footnote function work<span class="ref">
        <span class="refnum" onclick="footypop('test3')">[1]</span>
        <span id="test3" class="refbody">I think so...</span>
    </span>? Does it still work in the same paragraph? Let's try a reference<span class="ref">
        <span class="refnum" onclick="footypop('footyreftest')">[ref]</span>
        <span id="footyreftest" class="refbody">@MattDiss</span>
    </span>? Yup!


---
nocite: |
  @magrittr , @purrr , @xaringan , @data.table , @glmnet , @hqreg , @rlang , @msaenet , 
  @gridExtra , @grid , @ggplot2 , @magrittr
---

# References
