---
title: 'Demo: Custom K-Fold Subsetting Function'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
bibliography: ../References/Refs_Main.bib
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---

# Introduction, Preamble, and Setup {.unnumbered .tabset .tabset-fade .tabset-pills}

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

<style>
    .ref {
        position: relative;
        vertical-align: baseline;
    }

    .refnum {
        position: relative;
        left: -1px;
        bottom: 1ex;
        font-family: Verdana, sans-serif;
        color: #005994;
        font-size: .7em;
        font-weight: bold;
        text-decoration: underline;
        cursor: pointer;        
    }

    .refbody {
        font-family: Verdana, sans-serif;
        font-size: .7em;
        line-height: 1.1;
        display: none;
        min-width: 20em;
        position: absolute;
        left: 25px;
        bottom: 5px ;
        border: 1px solid;
        padding: 5px;
        background-color: #fff;
        word-wrap: break-word;
        z-index: 9999;
        overflow: auto;
    }

</style>

<script>
    function footypop(id) {
        var el = document.getElementById(id) ;
  
        if (el.style.display == "none") {
            el.style.display = "block" ;
        } else {
            el.style.display = "none" ;
        }
        
    }
</script>

## Introduction {.unnumbered}

This file demonstrates use of a custom function for indexing and splitting a full dataset for _k_-fold cross-validation. Use this procedure if you intend to utilize out-of-sample prediction error in some fashion for your application. I also use it here as an ad-hoc resampling procedure, since out-of-sample prediction error is not my primary metric of interest. Note that this does mean a reduction in effective sample size for any given sample. A bootstrap would be viable in this case, as well, and I hope to implement bootstramp resampling into this repo at some point in the future.


## Formatting Notes {.unnumbered}

This document makes use of **bolding**, _italics_, and "quotations" to help distinguish references to different types of items.

 * **bolding** will be used when referencing a new term or concept for the first time. Subsequent references to each term/concept will _not_ be bolded.
 * _italics_ will be used primarily for emphasis.
 * 'single quotations' will be used to clarify specific arguments for a function, or specific parameters of a mathemtical/statistical formulation
 * Inline references to code (functions, objects, specific commands, etc.) will use **<span style = "color:blue">`` `r "code_chunk_formatting"` ``</span>** in **<span style = "color:blue">`` `r "bolded blue font"` ``</span>**
 * Inline references to packages will similarly use *<span style = "color:red">`` `r "code_chunk_formatting"` ``</span>*, except in *<span style = "color:red">`` `r "italicized red font"` ``</span>*
 * References to other documents in this repository will use <span style = "color:green">`` `r "code_chunk_formatting"` ``</span> in <span style = "color:green">`` `r "un-italicized, un-bolded, green font"` ``</span>
 
The most important aspects of how I use spacing and line breaks are detailed below.

I generally put spaces between code inputs I consider "sufficiently distinct". This improves readability generally, but I find it particularly helpful for debugging. Note, however, that spaces are generally trivial in between distinct code elements in R, although this is not universally the case. Multi-character code elements, such as the pointer **<span style = "color:blue">`` `r "<-"` ``</span>** and most logical operators, _cannot_ include spaces in between components of the element. Note also that whitespace *is* meaningful in other programming languages, and so this convention should be considered with caution in your own practice.

Generally, I use line breaks to:

 * Break up separate arguments for a single command/function or chain of operations
 * To clearly distinguish between different closing parentheses, brackets, squigglies, etc., since RStudio will automatically tab each closing piece to align with its opening segment.


## Packages {.unnumbered}

Let's load the necessary packages. Links to more information about each package can be found in <a href="https://multach87.github.io/Robust_Lasso_ElasticNet/Component_Demos/00C_Robust_Issues.html"><span style = "color:green">`` `r "00B_Package_Descr_Refs"` ``</span></a>. Appropriate references for each package can be found in the "References" section at the end of this document.



```{r libraries}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
```


---
nocite: |
  @magrittr , @purrr , @xaringan , @mvtnorm , @data.table , @glmnet , @hqreg
---

## Data Loading {.unnumbered}

Let's load our demo data, which will just be our matricized singular applied dataset.

```{r load synthetic applied dataset}
demo.data <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/TestingApplied_Xmtx.RData")
```

Let's also quickly combine the X and Y list elements into a single object for convenience with the subsetting function. We will put Y into the first column since it corresponds with a left-to-right reading of a regression model.

```{r combine X and Y into single object}
demo.data2 <- setDT(data.frame(cbind(demo.data[["Y"]] , demo.data[["X"]]
                                     )
                               )
                    )
```

# Custom k-fold subsetting function {.tabset .tabset-fade .tabset-pills}

## Function Arguments

The chunk below takes a dataset and creates a numeric variable with *k* levels. The function takes as its arguments:

 * 'data' : a matrix/table/dataframe-type dataset to be split into *k* groups
 * 'y_col' : an index indicating the column number of the response variable Y in the original dataset
   * Defaults to the first column
 * 'x_cols' : indices indicating the column numbers of the predictor variables in the original dataset
   * Defaults to all but the first column
 * 'subset_col' : an index for the column number of the subset indicator in the intermediate dataset
   * Defaults to 1 + the number of columns in the original dataset
 * '*k'* : the number of folds for cross-validation
   * Defaults to *k = 5* folds
   * Corresponds with the sample size of the resulting holdout/testing set in the following manner:
   $$
   \begin{equation}
        n_{test} = \text{
                $n_{full}$ $/$ $k$
        }
   \end{equation}
   $$
 * 'seed' : a seed for fixing random processes
   * For this function, this is primarily associated with random and unordered assignment of the values of *k* across the full dataset
 * 'list' : this argument tells the function if you want the resulting *k*-fold indexed data to be put into list format, with each list element corresponding with one fold *k*
   * This function defaults to *FALSE*, meaning the resulting data will be in matrix/tabular format and will contain a separate column for the index *k*
   * If set to *TRUE*, the resulting list will *not* contain the index variable for *k*, as the list elements themselves correspond with *k*
   * If set to *"traintest"* (in quotes!), the resulting list will instead contain six elements:
     * "X" : The training set of predictor values ("X") containing predictor data from *k* - 1 folds/subsets
     * "Y" : The training set of response values
     * "X_Test" : The holdout/testing set of predictor values
     * "Y_Test" : The holdout/testing set of response values ("Y_Test")
     * "Seed" : The random-generating seed for the dataset. This is primarily for convenience with the multi-split wrapper function in <a href="https://multach87.github.io/Robust_Lasso_ElasticNet/Component_Demos/02B_KFold_Multi.html"><span style = "color:green">`` `r "02B_KFold_Multi"` ``</span></a>
     * "Subsets" : A *k*-level vector containing the row-ordered subset indicator for the full input dataset
     

## K-Fold Subsetting Function

```{r kfold subsetting function}
# k-fold subsetting function
kfold_subsetter <- function(data , y_col = 1 , 
                            x_cols = c(2:ncol(data)) , 
                            subset_col = (ncol(data) + 1) , 
                            k = 5 , seed = 7 , list = FALSE) {
        # check for string object in data argument
        if(is.character(data)) {
                data <- eval(parse_expr(data))
        }
        
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # check for 0 < k <= n/2
        if((k <= 0) | (k > (nrow(data) / 2) ) ) {
                stop("ERROR: number of folds 'k' must be greater than 0 and less than or equal to half of the sample size")
        }
        
        # determine number of subsets which contain an extra observation
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if 
        # # # # n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                    X = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset < k) %>%
                        .[ , c(1:length(x_cols)) , with = F] , #%>% 
                        #as.matrix() , 
                    Y = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset < k) %>%
                        .[ , 1 , with = F] , #%>%
                        #as.matrix() , 
                    X_Test = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset %in% k) %>%
                        .[ , c(1:length(x_cols)) , with = F] , # %>%
                        #as.matrix() , 
                    Y_Test = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset %in% k) %>%
                        .[ , 1 , with = F] , # %>%
                        #as.matrix() , 
                    Seed = seed , 
                    Subsets = newdata[ , "subset"]
                    )
                    )
        } else {
                newdata <- return(newdata)
        }
}



```

## Areas for Improvement

 * Take arguments for X/predictors and Y/outcome instead of a 'data' argument and 'x_cols' and 'y_cols'. These are idiosyncratic and not necessarily straightforward for a general user
 * The function itself can likely be streamlined for efficiency in a number of places. 

## Deprecated versions     
     
* The 08/05/2021 deprecation used *if* / *else if* branching to handle 1D/vector data vs. 2D/tabular data. The current version handles both data inputs without branching. 

```{r kfold subsetting function DEPRECATED 08052021 , eval = F}
# k-fold subsetting function
kfold_subsetter <- function(data , k = 5 , seed = 7 , list = FALSE) {
        if(length(dim(data)) == 2) { 
                # For 2D data input (matrices, data.tables, dataframes, etc.)
                
                # determine number of subsets which contain an extra element
                # # if n is not evenly divisible by k
                # # # note that this value will be 0 if n/k is evenly divisible
                nsams.large <- nrow(data) %% k
                
                # determine number of smaller subset if n 
                # # is not evenly divisible by k
                # # # note that this will be the total number of samples if n/k is evenly divisible
                nsams.small <- k - nsams.large
                
                # determine sample size of larger subsets if applicable
                samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
                
                # determine sample size of smaller/all subsets
                samsize.small <- floor(nrow(data) / k)
                
                # create indicator for each subset
                subset.indicator <- c(rep( (1 : k) , 
                                           floor(nrow(data) / k)
                                          ) ,
                                      rep( (1 : (nsams.large) ) , 
                                           (1 * (nsams.large != 0) ) 
                                          )
                                      )
                
                # fix random assignment process
                if(seed) {
                        set.seed(seed)
                }
                
                # combine subset indicator with original data  
                newdata <- cbind(data , 
                                 subset = sample(subset.indicator)
                                 )
                
                newdata <- setDT(data.frame(newdata))
                
                # create k-split list if desired
                if(list == TRUE) {
                        newdata <- return(split(newdata , 
                                                newdata[ , "subset"])
                                          )
                } else if(list == "TRAINTEST") {
                        newdata <- return(list(
                                Train = subset(newdata , subset < k) , 
                                Test = subset(newdata , subset %in% k)
                                )
                                )
                } else {
                        newdata <- return(newdata)
                }
        } else if (length(dim(data)) == 0){   
                # for 1D data
                
                # determine number of subsets which contain an extra element
                # # if n is not evenly divisible by k
                # # # note that this value will be 0 if n/k is evenly divisible
                nsams.large <- length(data) %% k
                
                # determine number of smaller subset if n 
                # # is not evenly divisible by k
                # # # note that this will be the total number of samples if n/k is evenly divisible
                nsams.small <- k - nsams.large
                
                # determine sample size of larger subsets if applicable
                samsize.large <- ceiling(length(data) / k) * (nsams.large != 0)
                
                # determine sample size of smaller/all subsets
                samsize.small <- floor(length(data) / k)
                
                # create indicator for each subset
                subset.indicator <- c(rep( (1 : k) , 
                                           floor(length(data) / k)
                                          ) ,
                                      rep( (1 : (nsams.large) ) , 
                                           (1 * (nsams.large != 0) ) 
                                          )
                                      )
                
                # fix random assignment process
                if(seed) {
                        set.seed(seed)
                }
                
                # combine original data and subset indicator
                newdata <- matrix(cbind(data , 
                                        subset = sample(subset.indicator) 
                                        ) ,
                                  ncol = 2
                                  )
                
                # setDT() for splitting step
                newdata <- setDT(data.frame(newdata))
                
                return(newdata)
                
                # create k-split list if desired
                if(list == TRUE) {
                        newdata <- return(split(newdata , 
                                                newdata[ , "subset"])
                                          )
                } else if(list == "TRAINTEST") {
                        newdata <- return(list(
                                Train = subset(newdata , subset < k) , 
                                Test = subset(newdata , subset %in% k)
                                )
                                )
                } else {
                        newdata <- return(newdata)
                }
        }
}
```

 * The 08/06/2021 deprecation only split the resulting data into training/testing/seed elements when setting the "list" argument to "traintest." The current version splits the data into the 6 elements described above.

```{r kfold subsetting function DEPRECATED2 08062021 , eval = F}
# k-fold subsetting function
kfold_subsetter <- function(data , k = 5 , seed = 7 , list = FALSE) {
        # check for string argument in "data"
        if(is.character(data)) {
                data <- eval(parse_expr(data))
        }
        
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # check for 0 < k <= n/2
        if((k <= 0) | (k > (nrow(data) / 2) ) ) {
                stop("ERROR: number of folds 'k' must be greater than 0 and less than sample size divided by 2")
        }
        
        # determine number of subsets which contain an extra element
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
                
        newdata <- setDT(data.frame(newdata))
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                        Train = subset(newdata , subset < k) , 
                        Test = subset(newdata , subset %in% k) , 
                        Seed = seed
                        )
                        )
        } else {
                newdata <- return(newdata)
        }
}
```

# Testing the function {.tabset .tabset-fade .tabset-pills}

Now let's test the function for all three list-splitting options.

## No Final Data Split

First, run the subsetter without splitting the final data.

```{r subset testing nonlist}
nonsplit.data <- kfold_subsetter(data = demo.data2 , k = 5 , 
                                 seed = 7 , list = FALSE)
```

And let's look at the structure of the data.

```{r nonlist str}
str(nonsplit.data)
```

The resulting object is a data.table/data.frame with 150 rows and 10 columns, with first column "Y" and last column "subset," and our potential predictors "X1" to "X8" in columns 2-9. There is also a row of information containing object attributes.

## Final List Containing *k* Elements

Next, run the subsetter but splitting by fold index *k*.

```{r subset testing full split}
fullsplit.data <- kfold_subsetter(data = demo.data2 , k = 5 , 
                                  seed = 7 , list = TRUE)
```

And let's look at the structure of the data.

```{r full split str}
str(fullsplit.data)
```

**<span style = "color:blue">`` `r "str()"` ``</span>** now produces much more information than before. Now there is an additional layer of information corresponding with the list structure, in addition to the data.table/.frame layer of information similar to what we saw previously.

Let's run that chunk again. This time, though, we're going to specify an additional argument to **<span style = "color:blue">`` `r "str()"` ``</span>**:

```{r full split str lvl 1}
str(fullsplit.data , max.level = 1)
```

Now, we only see the list-level structure of the object and basic information about the data.frames/.tables within each list element. Setting 'max.level = 2' would simply produce the same structural output as **<span style = "color:blue">`` `r "str(fullsplit.data)"` ``</span>**. specifying larger values for 'max.level' would not provide any additional information for the current context, as **<span style = "color:blue">`` `r "fullsplit.data"` ``</span>** only has two levels.

## Final List Containing Training and Testing Elements

Finally, run the subsetter but splitting by a single training/testing split.

```{r subset testing traintest split}
trtstsplit.data <- kfold_subsetter(data = demo.data2 , k = 5 , 
                                  seed = 7 , list = "traintest")
```

And let's look at the structure of the data.

```{r traintest split str}
str(trtstsplit.data)
```

This last specification for **<span style = "color:blue">`` `r "kfold_subsetter()"` ``</span>** produces another list, except now we see separate list elements for the training set of our potential predictors X1-X8, the training set of our response Y, the holdout/testing set of our predictors, the holdout/testint set of our response, our random-fixing seed, and our subset indices for each row in the original dataset. All of the list elements are data.tables/.frames except the seed, which is just a single-element numeric vector.

Let's specify 'max.level = 1' for this context and see what is produced, and compare it to the 1st-level structure of **<span style = "color:blue">`` `r "fullsplit.data"` ``</span>**

```{r traintest split str level 1}
str(trtstsplit.data , max.level = 1)
```

```{r fullsplit str level 1 comparison}
str(fullsplit.data , max.level = 1)
```


# References {.unnumbered}
