---
title: 'Demo: Producting Multiple Cross-Validation Splits'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
bibliography: ../References/Refs_Main.bib
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---



# Introduction, Preamble, and Setup {.unnumbered .tabset .tabset-fade .tabset-pills}

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

<style>
    .ref {
        position: relative;
        vertical-align: baseline;
    }

    .refnum {
        position: relative;
        left: -1px;
        bottom: 1ex;
        font-family: Verdana, sans-serif;
        color: #005994;
        font-size: .7em;
        font-weight: bold;
        text-decoration: underline;
        cursor: pointer;        
    }

    .refbody {
        font-family: Verdana, sans-serif;
        font-size: .7em;
        line-height: 1.1;
        display: none;
        min-width: 20em;
        position: absolute;
        left: 25px;
        bottom: 5px ;
        border: 1px solid;
        padding: 5px;
        background-color: #fff;
        word-wrap: break-word;
        z-index: 9999;
        overflow: auto;
    }

</style>

<script>
    function footypop(id) {
        var el = document.getElementById(id) ;
  
        if (el.style.display == "none") {
            el.style.display = "block" ;
        } else {
            el.style.display = "none" ;
        }
        
    }
</script>

## Introduction {.unnumbered}

This file extends the *k*-fold subsetting function for conducting multiple training-testing splits of a single dataset.

## Formatting Notes {.unnumbered}

This document makes use of **bolding**, _italics_, and "quotations" to help distinguish references to different types of items.

 * **bolding** will be used when referencing a new term or concept for the first time. Subsequent references to each term/concept will _not_ be bolded.
 * _italics_ will be used primarily for emphasis.
 * 'single quotations' will be used to clarify specific arguments for a function, or specific parameters of a mathemtical/statistical formulation
 * Inline references to code (functions, objects, specific commands, etc.) will use **<span style = "color:blue">`` `r "code_chunk_formatting"` ``</span>** in **<span style = "color:blue">`` `r "bolded blue font"` ``</span>**
 * Inline references to packages will similarly use *<span style = "color:red">`` `r "code_chunk_formatting"` ``</span>*, except in *<span style = "color:red">`` `r "italicized red font"` ``</span>*
 * References to other documents in this repository will use <span style = "color:green">`` `r "code_chunk_formatting"` ``</span> in <span style = "color:green">`` `r "un-italicized, un-bolded, green font"` ``</span>
 
The most important aspects of how I use spacing and line breaks are detailed below.

I generally put spaces between code inputs I consider "sufficiently distinct". This improves readability generally, but I find it particularly helpful for debugging. Note, however, that spaces are generally trivial in between distinct code elements in R, although this is not universally the case. Multi-character code elements, such as the pointer **<span style = "color:blue">`` `r "<-"` ``</span>** and most logical operators, _cannot_ include spaces in between components of the element. Note also that whitespace *is* meaningful in other programming languages, and so this convention should be considered with caution in your own practice.

Generally, I use line breaks to:

 * Break up separate arguments for a single command/function or chain of operations
 * To clearly distinguish between different closing parentheses, brackets, squigglies, etc., since RStudio will automatically tab each closing piece to align with its opening segment.


## Packages {.unnumbered}

Let's load the necessary packages. Links to more information about each package can be found in <a href="https://multach87.github.io/Robust_Lasso_ElasticNet/Component_Demos/00C_Robust_Issues.html"><span style = "color:green">`` `r "00B_Package_Descr_Refs"` ``</span></a>. Appropriate references for each package can be found in the "References" section at the end of this document.



```{r libraries}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
library(rlang)      # For parse_expr() to parse data name from function argument
```


---
nocite: |
  @magrittr , @purrr , @xaringan , @mvtnorm , @data.table , @glmnet , @hqreg , @rlang
---

## Data Loading {.unnumbered}

Let's load our demo data, which will just be our matricized singular applied dataset.

```{r load synthetic applied dataset}
demo.data <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/TestingApplied_Xmtx.RData")
```

Let's also quickly combine the X and Y list elements into a single object for convenience with the subsetting function. We will put Y into the first column since it corresponds with a left-to-right reading of a regression model.

```{r combine X and Y into single object}
demo.data2 <- setDT(data.frame(cbind(demo.data[["Y"]] , demo.data[["X"]]
                                     )
                               )
                    )
```

# Custom k-fold subsetting function {.tabset .tabset-fade .tabset-pills}

## Original Function

This function is as described in the <a href="https://multach87.github.io/Robust_Lasso_ElasticNet/Component_Demos/02A_KFold_Subsetter.html"><span style = "color:green">`` `r "02A_KFold_Subsetter"` ``</span></a>. A full description of the arguments of the original function can be found therein.

```{r kfold subsetting function , echo = F}
# k-fold subsetting function
kfold_subsetter <- function(data , y_col = 1 , 
                            x_cols = c(2:ncol(data)) , 
                            subset_col = (ncol(data) + 1) , 
                            k = 5 , seed = 7 , list = FALSE) {
        # check for string object in data argument
        if(is.character(data)) {
                data <- eval(parse_expr(data))
        }
        
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # check for 0 < k <= n/2
        if((k <= 0) | (k > (nrow(data) / 2) ) ) {
                stop("ERROR: number of folds 'k' must be greater than 0 and less than or equal to half of the sample size")
        }
        
        # determine number of subsets which contain an extra observation
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if 
        # # # # n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                    X = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset < k) %>%
                        .[ , c(1:length(x_cols)) , with = F] , #%>% 
                        #as.matrix() , 
                    Y = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset < k) %>%
                        .[ , 1 , with = F] , #%>%
                        #as.matrix() , 
                    X_Test = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset %in% k) %>%
                        .[ , c(1:length(x_cols)) , with = F] , # %>%
                        #as.matrix() , 
                    Y_Test = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset %in% k) %>%
                        .[ , 1 , with = F] , # %>%
                        #as.matrix() , 
                    Seed = seed , 
                    Subsets = newdata[ , "subset"]
                    )
                    )
        } else {
                newdata <- return(newdata)
        }
}



```


## Multi-Subsetting Wrapper

The following function serves as a wrapper around the original subsetting function which incorporates additional arguments. These arguments extend the original function to conduct multiple training-testing splits and store the resulting datasets in a singular overarching list.

The wrapper's arguments are as follows. The ellipses are arguments for the original subsetting function which are handled internally within the wrapper.

 * 'data' : a matrix/data.table/data.frame-type data object to be split into training-testing sets
 * 'seed_multi' : a seed for generating the vector of seeds for each individual training-testing *k*-fold split
 * 'num_splits' : the number of training-testing splits to be created
    * Defaults to 100
 * 'test_percent' : the percentage (as a decimal) of the full data to be reserved for the holdout/testing set
    * Corresponds with the number of folds *k* for *k*-fold subsetting in the following manner:
   $$
   \begin{equation}
        test\_percent = \text{
                $1$ $/$ $k$
        }
   \end{equation}
   $$
    * Defaults to 20$\%$, which corresponds with 5-fold subsetting

Note that this wrapper defaults to setting 'list = "traintest"' in **<span style = "color:blue">`` `r "kfold_subsetter()"` ``</span>**. This decision was implemented to reduce issues with formatting an applied dataset for the final combined-model function. The subsetting function is relatively more flexible in terms of data inputs than the full combined-model function, therefore I wanted to streamline data formatting from that point forward when using the combined model in full.

```{r multi-kfold wrapper}
kfold_multi <- function(data , ... , seed_multi = 713 , 
                        num_splits = 100 , test_percent = .2) {
        # set the seed for generating individual kfold seeds
        set.seed(seed_multi)
        
        # turn data object name into character string for subsequent use
        data.name <- deparse(substitute(data))
        
        # generate random numbers equal to num_splits so that each
        # # individual run of the subsetter has a unique seed
        seeds <- sample(c(1:100000) , size = num_splits , replace = FALSE)
        
        # set value of k corresponding with test_percent
        if((test_percent >= 1) | (test_percent <= 0)) {
                stop("ERROR: 'test_percent' must be greater than and less than 1")
        }
        
        # set k for individual splits
        k <- 1 / test_percent
        
        # initialize and data.table of kfold arguments
        split_repped.dt <- setDT(as.data.frame(matrix(ncol = 1 , 
                                              nrow = num_splits)))
        
        # fill data.table with arguments for kfold function
        split_repped.dt[ , ':=' (data = data.name , 
                         k = k , 
                         seed = seeds , 
                         list = "traintest"
                         )
                 ]
        
        # remove blank column from initializing
        split_repped.dt <- split_repped.dt[ , !1 , with = F]
        
        # run subsetter for the desired number of splits
        full.data <- split_repped.dt %>%
                pmap(kfold_subsetter)
        
        # store single object of all training/testing splits
        return(full.data)
}
```


# Testing the Function {.tabset .tabset-fade .tabset-pills}

## Testing

Let's do an example run of the multi-subsetting wrapper function. To improve readability of the **<span style = "color:blue">`` `r "str()"` ``</span>** on the resulting object, I'm going to set 'num_splits = 5'.

```{r multi-kfold testing}
trntst_multi.data <- kfold_multi(data = demo.data2 , 
                                 num_splits = 5)
```

## Resulting Object Structure

Let's check out the structure of the resulting object.

```{r str trntst multi}
str(trntst_multi.data)
```

The output from **<span style = "color:blue">`` `r "str()"` ``</span>** shows three levels: 
 * 1st Level: A list, with each element corresponding with a different k-fold split of the singular data input

And 2nd/3rd levels corresponding with the result of running **<span style = "color:blue">`` `r "kfold_subsetter(demo.data2 , ... , list = \"traintest\")"` ``</span>** on a single dataset.

And we could run the **<span style = "color:blue">`` `r "str()"` ``</span>** command with 'max.level = 2' and 'max.level = 1'

```{r str trntst multi lvl 2}
str(trntst_multi.data , max.level = 2)
```

```{r str trntst multi lvl 1}
str(trntst_multi.data , max.level = 1)
```

# References