---
title: 'Process Coefficients from a Single Adaptation and All Data Splits'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: false
bibliography: Component_Refs.bib
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---

## Introduction

This document demonstrates the processing of model coefficients from a list of multiple model objects produced by the previous **<span style = "color:blue">`` `r "hqmsa.sim.funct()"` ``</span>** function. Note that the function created herein will be designed to run on all model results from a single adaptation, _across all data splits_. 

## Preamble and Setup

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

### A note on formatting

This document makes use of **bolding**, _italics_, and "quotations" to help distinguish different types of items being referenced.

 * **bolding** will be used when referencing a new term or concept for the first time. Subsequent references to each term/concept will _not_ be bolded.
 * _italics_ will be used primarily for emphasis.
 * 'single quotations' will be used to clarify specific arguments for a function, or specific parameters of a mathemtical/statistical formulation
 * Inline references to code (functions, objects, specific commands, etc.) will use **<span style = "color:blue">`` `r "code_chunk_formatting"` ``</span>** in **<span style = "color:blue">`` `r "bolded blue font"` ``</span>**
 * Inline references to packages will similarly use *<span style = "color:red">`` `r "code_chunk_formatting"` ``</span>*, except in *<span style = "color:red">`` `r "italicized red font"` ``</span>*
 
Re: spacing and line breaks - I'm pretty heterogeneous in my application of line breaks and spacing, in a way that is idiosyncratic to my own code practice. The most important aspects of my spacing and line breaks are detailed below

I generally put spaces between code inputs I consider "sufficiently distinct". This improves readability generally, but I find it particularly helpful for debugging. Note, however, that spaces are generally trivial in between distinct code inputs in R, although this is not universally the case. Multi-character code inputs, such as the pointer **<span style = "color:blue">`` `r "<-"` ``</span>** and most logical operators, _cannot_ include spaces in between components of the code input. Note also that whitespace *is* meaningful in other programming languages, and so this convention should be considered with caution in your own practice.

Generally, I use line breaks to:

 * Break up separate arguments for a single command/function or chain of operations
 * To clearly distinguish between different closing parentheses, brackets, squigglies, etc., since RStudio will automatically tab each closing piece to match its opener.
 
### Packages

Let's load the necessary packages. Links to more information about each packages and helpful guides (where applicable) can be found in 00B_Package_Descr_Refs. Appropriate references for each package can be found in the "References" section at the end of this document.

```{r libraries , warning = FALSE}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
library(rlang)      # For parse_expr() to parse data name in k-fold subsetting functions
library(msaenet)    # For the multi-step adaptive elastic net
```

---
nocite: |
  @magrittr , @purrr , @xaringan , @mvtnorm , @data.table , @glmnet , @hqreg , @rlang , @msaenet
---

### Load and Process 10-Split Model Data

Let's also load our model data. Note that this is a collection of all adaptations run 

```{r load model results}
models_all <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/models10.RData")
```

Let's also store the results from a single adaptation for the purposes of creating and demo'ing the function. We're going to use the results from the adaptive LAD lasso.

```{r store ladlasso}
ladlasso.models <- models_all[["ladlasso"]]
```


And finally let's get rid of the full model object from the environment.

```{r remove combined model data}
rm(list = c("models_all"))
```

## Extracting Coefficient Information

### Coefficient frequency function

There are many pieces of information that might be extracted from such a rich collection of model results. Let's focus, however, on predictor selection/inclusion frequency.

I'd like to add a list element that contains information about the adaptive LAD lasso models across all data splits, particularly with respect to coefficient inclusion frequency. Note that this does _not_ include intercepts, given that some models do not include them.

Also note that this function relies on a list with a comparable structure to my processed models:
 
 * Level 1 of the list corresponds with each data split
 * Level 2 of the list corresponds with different model result objects from a given data split
   * One list element at Level 2 contains the model summary object 'model.info'
     * which contains a list element 'coefs', a numeric vector of coefficient values (and intercept, if applicable)

```{r extract coefficient information}
coefs_freq <- function(models) {
       # initialize vector of predictor selection frequency, with all values set to 0
       coefs.freq <- numeric(length = 8)
       
       # label frequency vector with predictor names, accounting for 
       names(coefs.freq) <- names(models[[1]][["model.info"]][["coefs"]][-1])
       
       # update each predictors' selection frequency if the corresponding coefficient value
       # # in each data split is nonzero (aka, was selected into the model)
       for(i in 1:length(models)) {
              for(j in 1:length(coefs.freq)) {
                     if(models[[i]][["model.info"]][["coefs"]][(j + 1)] != 0) {
                            coefs.freq[j] <- coefs.freq[j] + 1
                     }
              }
       }
       
       return(coefs.freq)       
}
```

## Example 

Now let's run it and see what happens when I run the function.

```{r run coef frequency function on ladlasso models}
ladlasso.models[["coef.freq"]] <- ladlasso.models %>%
       coefs_freq
```

And let's examine the structure of the resulting frequency object within **<span style = "color:blue">`` `r "ladlasso.models"` ``</span>**. The other elements of the model object will resemble what was seen in 03A-C.

```{r coeffreq str}
str(ladlasso.models[["coef.freq"]])
```

## References