---
title: 'Demo: Splitting Data for Cross-Validation and Improved Model Results'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: html_document
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---

## Introduction

This file demonstrates extends the *k*-fold subsetting function for conducting multiple training-testing splits of a single dataset.

## Preamble

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

First, let's load the necessary packages.

```{r libraries}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
library(tibble)
```

Let's also load our demo data, which will just be our matricized singular applied dataset.

```{r load synthetic applied dataset}
demo.data <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/TestingApplied_Xmtx.RData")
```

Let's also quickly combine the X and Y list elements into a single object for convenience with the subsetting function. We will put Y into the first column since it corresponds with a left-to-right reading of a regression model.

```{r combine X and Y into single object}
demo.data2 <- setDT(data.frame(cbind(demo.data[["Y"]] , demo.data[["X"]]
                                     )
                               )
                    )
```

## Custom k-fold subsetting function

This function is as described in the "KFold_Subsetter" file. A full description of the arguments of the original function can be found therein.

The chunk below takes a dataset and creates an ordinal variable with *k* levels, corresponding with each of the *k* test sets. The function takes as its arguments:

 * **data** : a dataset to be split into *k* groups
 * ***k*** : the number of folds for cross-validation
   * Defaults to 5 folds
   * Corresponds with the percentage of data to be used as a holdout/testing set in the following manner:
   $$
   \begin{equation}
        n_{test} = \text{
                $n_{full}$ $/$ $k$
        }
   \end{equation}
   $$
 * **seed** : a seed for fixing random processes
   * For this function, this is primarily associated with random and unordered assignment of the values of *k* across the full dataset
 * **list** : this argument tells the function if you want the resulting *k*-fold indexed data to be put into list format, with each list element corresponding with one fold *k*
   * This function defaults to *FALSE*, meaning the resulting data will be in matrix/tabular format and will contain a separate column for the index *k*
   * If set to *TRUE*, the resulting list will *not* contain the index variable for *k*, as the list elements themselves correspond with *k*
   * If set to *"traintest"* (in quotes!), the resulting list will instead contain two elements:
     * A training set containing data from *k* - 1 folds/subsets
     * A holdout/testing set containing data from the *k* fold/subset
     
Although the chunk isn't printed in the .html walkthrough, the .Rmd file contains an additional chunk of a deprecated version of this function. The deprecated version used *if* / *else if* branching to handle 1D/vector data vs. 2D/tabular data. The current version handles both data inputs without branching.

```{r kfold subsetting function}
# k-fold subsetting function
kfold_subsetter <- function(data , k = 5 , seed = 7 , list = FALSE) {
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # determine number of subsets which contain an extra element
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
                
        newdata <- setDT(data.frame(newdata))
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                        Train = subset(newdata , subset < k) , 
                        Test = subset(newdata , subset %in% k)
                        )
                        )
        } else {
                newdata <- return(newdata)
        }
}
```


## Multi-Subsetting Wrapper

The following function serves as a wrapper around the original subsetting function which incorporates additional arguments. These arguments extend the original function to conduct multiple training-testing splits and store the resulting datasets in a singular overarching list.

```{r multi-kfold wrapper}

```