---
title: 'Demo: Splitting Data for Cross-Validation and Improved Model Results'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: html_document
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---

## Introduction

This file demonstrates extends the *k*-fold subsetting function for conducting multiple training-testing splits of a single dataset.

## Preamble

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

First, let's load the necessary packages.

```{r libraries}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
library(rlang)      # For parse_expr() to parse data name from argument
```

Let's also load our demo data, which will just be our matricized singular applied dataset.

```{r load synthetic applied dataset}
demo.data <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/TestingApplied_Xmtx.RData")
```

Let's also quickly combine the X and Y list elements into a single object for convenience with the subsetting function. We will put Y into the first column since it corresponds with a left-to-right reading of a regression model.

```{r combine X and Y into single object}
demo.data2 <- setDT(data.frame(cbind(demo.data[["Y"]] , demo.data[["X"]]
                                     )
                               )
                    )
```

## Custom k-fold subsetting function

This function is as described in the "KFold_Subsetter" file. A full description of the arguments of the original function can be found therein.

```{r kfold subsetting function}
# k-fold subsetting function
kfold_subsetter <- function(data , y_col = 1 , 
                            x_cols = c(2:ncol(data)) , 
                            subset_col = (ncol(data) + 1) , 
                            k = 5 , seed = 7 , list = FALSE) {
        # check for string object in data argument
        if(is.character(data)) {
                data <- eval(parse_expr(data))
        }
        
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # check for 0 < k <= n/2
        if((k <= 0) | (k > (nrow(data) / 2) ) ) {
                stop("ERROR: number of folds 'k' must be greater than 0 and less than sample size divided by 2")
        }
        
        # determine number of subsets which contain an extra element
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
                
        newdata <- setDT(data.frame(newdata))
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                    X = subset(newdata[ , c(x_cols, subset_col) , with = F] , 
                                         subset < k) %>%
                        .[ , c(1:8) , with = F] %>%
                        as.matrix() , 
                    Y = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset < k) %>%
                        .[ , 1 , with = F] %>%
                        as.matrix() , 
                    X_Test = subset(newdata[ , c(x_cols, subset_col) , 
                                            with = F] , subset %in% k) %>%
                        .[ , c(1:8) , with = F] %>%
                        as.matrix() , 
                    Y_Test = subset(newdata[ , c(y_col , subset_col) , 
                                        with = F] , subset %in% k) %>%
                        .[ , 1 , with = F] %>%
                        as.matrix() , 
                    Seed = seed , 
                    Subsets = newdata[ , "subset"]
                        #Train = subset(newdata , subset < k) , 
                        #Test = subset(newdata , subset %in% k) , 
                        #Seed = seed
                        )
                        )
        } else {
                newdata <- return(newdata)
        }
}
```


## Multi-Subsetting Wrapper

The following function serves as a wrapper around the original subsetting function which incorporates additional arguments. These arguments extend the original function to conduct multiple training-testing splits and store the resulting datasets in a singular overarching list.

The wrapper's arguments are as follows. The ellipses are arguments for the original subsetting function which are handled internally within the wrapper.

 * **data** : a dataset to be split into training-testing sets
 * **seed_multi** : a seed for generating the vector of seeds for each individual training-testing *k*-fold split
 * **num_splits** : the number of training-testing splits to be created
    * Defaults to 100
 * **test_percent** : the percentage (as a decimal) of the full data to be reserved for the holdout/testing set
    * Corresponds with the number of folds *k* for *k*-fold subsetting in the following manner:
   $$
   \begin{equation}
        test\_percent = \text{
                $1$ $/$ $k$
        }
   \end{equation}
   $$
    * Defaults to 20$\%$, which corresponds with 5-fold subsetting

Note that this wrapper defaults to setting "list" argument in *kfold_subsetter()* to "traintest." This decision was implemented to reduce issues with formatting an applied dataset for the final combined-model function. The subsetting function is relatively more flexible in terms of data inputs than the full combined-model function, therefore I wanted to streamline data formatting from that point forward when using the combined model in full.

```{r multi-kfold wrapper}
kfold_multi <- function(data , ... , seed_multi = 713 , 
                        num_splits = 100 , test_percent = .2) {
        # set the seed for generating individual kfold seeds
        set.seed(seed_multi)
        
        # turn data object name into character string for subsequent use
        data.name <- deparse(substitute(data))
        
        # generate random numbers equal to num_splits so that each
        # # individual run of the subsetter has a unique seed
        seeds <- sample(c(1:100000) , size = num_splits , replace = FALSE)
        
        # set value of k corresponding with test_percent
        if((test_percent >= 1) | (test_percent <= 0)) {
                stop("ERROR: 'test_percent' must be greater than and less than 1")
        }
        
        # set k for individual splits
        k <- 1 / test_percent
        
        # initialize and data.table of kfold arguments
        split_repped.dt <- setDT(as.data.frame(matrix(ncol = 1 , 
                                              nrow = num_splits)))
        
        # fill data.table with arguments for kfold function
        split_repped.dt[ , ':=' (data = data.name , 
                         k = k , 
                         seed = seeds , 
                         list = "traintest"
                         )
                 ]
        
        # remove blank column from initializing
        split_repped.dt <- split_repped.dt[ , !1 , with = F]
        
        # run subsetter for the desired number of splits
        full.data <- split_repped.dt %>%
                pmap(kfold_subsetter)
        
        # store single object of all training/testing splits
        return(full.data)
}
```


## Examples

INCLUDE SOME EXAMPLES