---
title: 'K-Fold Subsetting Function Demo'
author: "Matt Multach"
date: "`r Sys.Date()`"
output: html_document
#output: 
#       xaringan::infinite_moon_reader:
#              lib_dir: lib
#              highlightStyle: github
#              highlightLines: true
#              countIncrementalSlides: false
#              beforeInit: "macros.js"
#              css: [default, tamu, tamu-fonts]
---

## Introduction

This file demonstrates use of a custom function for indexing and splitting a full dataset for k-fold cross-validation.

## Preamble

```{r setup, message = FALSE , warning = FALSE , include=FALSE}
knitr::opts_chunk$set(echo = TRUE , eval = TRUE)
```

First, let's load the necessary packages.

```{r libraries}
# This chunk loads the packages used in this workbook
library(xaringan)   # Allows active preview of report in RStudio
library(mvtnorm)    # Generates multivariate-normal data
library(magrittr)   # Used for piping
library(purrr)      # Used for mapping functions efficiently
library(data.table) # For more streamlined data structures
library(glmnet)     # For general lasso/elastic net functionality
library(hqreg)      # For LAD/Huber-loss regularization with SNCD algorithm
```


Let's also load our demo data, which will just be our matricized singular applied dataset.

```{r load synthetic applied dataset}
demo.data <- readRDS("/Users/Matt/Desktop/GitHub_Dissertation_Scripts/Robust_Lasso_ElasticNet/Datasets/TestingApplied_Xmtx.RData")
```

Let's also quickly combine the X and Y list elements into a single object for convenience with the subsetting function. We will put Y into the first column since it corresponds with a left-to-right reading of a regression model.

```{r combine X and Y into single object}
demo.data2 <- setDT(data.frame(cbind(demo.data[["Y"]] , demo.data[["X"]]
                                     )
                               )
                    )
```

## Custom k-fold subsetting function

The chunk below takes a dataset and creates an ordinal variable with *k* levels, corresponding with each of the *k* test sets. The function takes as its arguments:

 * **data** : a dataset to be split into *k* groups
 * ***k*** : the number of folds for cross-validation
   * Defaults to 5 folds
   * Corresponds with the percentage of data to be used as a holdout/testing set in the following manner:
   $$
   \begin{equation}
        n_{test} = \text{
                $n_{full}$ $/$ $k$
        }
   \end{equation}
   $$
 * **seed** : a seed for fixing random processes
   * For this function, this is primarily associated with random and unordered assignment of the values of *k* across the full dataset
 * **list** : this argument tells the function if you want the resulting *k*-fold indexed data to be put into list format, with each list element corresponding with one fold *k*
   * This function defaults to *FALSE*, meaning the resulting data will be in matrix/tabular format and will contain a separate column for the index *k*
   * If set to *TRUE*, the resulting list will *not* contain the index variable for *k*, as the list elements themselves correspond with *k*
   * If set to *"traintest"* (in quotes!), the resulting list will instead contain three elements:
     * A training set containing data from *k* - 1 folds/subsets
     * A holdout/testing set containing data from the *k* fold/subset
     * The random-generating seed for the dataset. This is primarily for convenience with the multi-split wrapper function in "CV_Splitting"
     
Although the chunk isn't printed in the .html walkthrough, the .Rmd file contains an additional chunk of a deprecated version of this function. The deprecated version used *if* / *else if* branching to handle 1D/vector data vs. 2D/tabular data. The current version handles both data inputs without branching.

```{r kfold subsetting function DEPRECATED , echo = F , eval = F}
# k-fold subsetting function
kfold_subsetter <- function(data , k = 5 , seed = 7 , list = FALSE) {
        if(length(dim(data)) == 2) { 
                # For 2D data input (matrices, data.tables, dataframes, etc.)
                
                # determine number of subsets which contain an extra element
                # # if n is not evenly divisible by k
                # # # note that this value will be 0 if n/k is evenly divisible
                nsams.large <- nrow(data) %% k
                
                # determine number of smaller subset if n 
                # # is not evenly divisible by k
                # # # note that this will be the total number of samples if n/k is evenly divisible
                nsams.small <- k - nsams.large
                
                # determine sample size of larger subsets if applicable
                samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
                
                # determine sample size of smaller/all subsets
                samsize.small <- floor(nrow(data) / k)
                
                # create indicator for each subset
                subset.indicator <- c(rep( (1 : k) , 
                                           floor(nrow(data) / k)
                                          ) ,
                                      rep( (1 : (nsams.large) ) , 
                                           (1 * (nsams.large != 0) ) 
                                          )
                                      )
                
                # fix random assignment process
                if(seed) {
                        set.seed(seed)
                }
                
                # combine subset indicator with original data  
                newdata <- cbind(data , 
                                 subset = sample(subset.indicator)
                                 )
                
                newdata <- setDT(data.frame(newdata))
                
                # create k-split list if desired
                if(list == TRUE) {
                        newdata <- return(split(newdata , 
                                                newdata[ , "subset"])
                                          )
                } else if(list == "TRAINTEST") {
                        newdata <- return(list(
                                Train = subset(newdata , subset < k) , 
                                Test = subset(newdata , subset %in% k)
                                )
                                )
                } else {
                        newdata <- return(newdata)
                }
        } else if (length(dim(data)) == 0){   
                # for 1D data
                
                # determine number of subsets which contain an extra element
                # # if n is not evenly divisible by k
                # # # note that this value will be 0 if n/k is evenly divisible
                nsams.large <- length(data) %% k
                
                # determine number of smaller subset if n 
                # # is not evenly divisible by k
                # # # note that this will be the total number of samples if n/k is evenly divisible
                nsams.small <- k - nsams.large
                
                # determine sample size of larger subsets if applicable
                samsize.large <- ceiling(length(data) / k) * (nsams.large != 0)
                
                # determine sample size of smaller/all subsets
                samsize.small <- floor(length(data) / k)
                
                # create indicator for each subset
                subset.indicator <- c(rep( (1 : k) , 
                                           floor(length(data) / k)
                                          ) ,
                                      rep( (1 : (nsams.large) ) , 
                                           (1 * (nsams.large != 0) ) 
                                          )
                                      )
                
                # fix random assignment process
                if(seed) {
                        set.seed(seed)
                }
                
                # combine original data and subset indicator
                newdata <- matrix(cbind(data , 
                                        subset = sample(subset.indicator) 
                                        ) ,
                                  ncol = 2
                                  )
                
                # setDT() for splitting step
                newdata <- setDT(data.frame(newdata))
                
                return(newdata)
                
                # create k-split list if desired
                if(list == TRUE) {
                        newdata <- return(split(newdata , 
                                                newdata[ , "subset"])
                                          )
                } else if(list == "TRAINTEST") {
                        newdata <- return(list(
                                Train = subset(newdata , subset < k) , 
                                Test = subset(newdata , subset %in% k)
                                )
                                )
                } else {
                        newdata <- return(newdata)
                }
        }
}
```

```{r kfold subsetting function}
# k-fold subsetting function
kfold_subsetter <- function(data , k = 5 , seed = 7 , list = FALSE) {
        # check for string argument in "data"
        if(is.character(data)) {
                data <- eval(parse_expr(data))
        }
        
        # check for data.table and setDT() if not a data.table
        if(!(TRUE %in% (class(data) == "data.table"))) {
                data <- setDT(data.frame(data))
        }
        
        # check for 0 < k <= n/2
        if((k <= 0) | (k > (nrow(data) / 2) ) ) {
                stop("ERROR: number of folds 'k' must be greater than 0 and less than sample size divided by 2")
        }
        
        # determine number of subsets which contain an extra element
        # # if n is not evenly divisible by k
        # # # note that this value will be 0 if n/k is evenly divisible
        nsams.large <- nrow(data) %% k
        
        # determine number of smaller subset if n 
        # # is not evenly divisible by k
        # # # note that this will be the total number of samples if n/k is evenly divisible
        nsams.small <- k - nsams.large
        
        # determine sample size of larger subsets if applicable
        samsize.large <- ceiling(nrow(data) / k) * (nsams.large != 0)
        
        # determine sample size of smaller/all subsets
        samsize.small <- floor(nrow(data) / k)
                
        # create indicator for each subset
        subset.indicator <- c(rep( (1 : k) , 
                                   floor(nrow(data) / k)
                                  ) ,
                              rep( (1 : (nsams.large) ) , 
                                   (1 * (nsams.large != 0) ) 
                                  )
                              )
                
        # fix random assignment process
        if(seed) {
                set.seed(seed)
        }
        
        # combine subset indicator with original data  
        newdata <- cbind(data , 
                         subset = sample(subset.indicator)
                         )
                
        newdata <- setDT(data.frame(newdata))
        
        # create k-split list if desired
        if(list == TRUE) {
                newdata <- return(split(newdata , 
                                        newdata[ , "subset"])
                                  )
        } else if(list == "traintest") {
                newdata <- return(list(
                        Train = subset(newdata , subset < k) , 
                        Test = subset(newdata , subset %in% k) , 
                        Seed = seed
                        )
                        )
        } else {
                newdata <- return(newdata)
        }
}
```

## Testing the function

Now let's test the function for all three list-splitting options.

### K-fold subset testing: no final data split

First, run the subsetter without splitting the final data.

```{r subset testing nonlist}
nonsplit.data <- kfold_subsetter(data = demo.data2 , k = 5 , 
                                 seed = 7 , list = FALSE)
```

And let's look at the structure of the data.

```{r nonlist str}
str(nonsplit.data)
```

### K-fold subset testing: list containing *k* elements

Next, run the subsetter but splitting by fold index *k*.

```{r subset testing full split}
fullsplit.data <- kfold_subsetter(data = demo.data2 , k = 5 , 
                                  seed = 7 , list = TRUE)
```

And let's look at the structure of the data.

```{r full split str}
str(fullsplit.data)
```

### K-fold subset testing: list containing training and testing elements

Next, run the subsetter but splitting by a single training/testing split.

```{r subset testing traintest split}
trtstsplit.data <- kfold_subsetter(data = demo.data2 , k = 5 , 
                                  seed = 7 , list = "traintest")
```

And let's look at the structure of the data.

```{r traintest split str}
str(trtstsplit.data)
```
